Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-22 paper/main.tex
More slicing information

\subsubsection{Slice (Dynamic)}
\begin{figure}
\begin{lstlisting}
x = input() # The user inputs 0
if x != 0:
  x = 1
print("One over your number is: %d" % (1 / x))
\end{lstlisting}
\caption{In this program, the programmer may have intended the placeholder
  assignment \lt{x = 1} to be used in case the user input 0, and used the
  incorrect condition \lt{x != 0}, rather than \lt{x == 0}. However, the
  condition expression will not be included in the slice (which would include
  lines 1 and 4), because the presence of
  the conditional does not affect the behavior of the program.
}
\label{slice-downside-example}
\end{figure}

\begin{figure}
\begin{lstlisting}
while true:
  x = input()
  if x != 0:
    break
  print("One over your number is: %d" % (1 / x))
\end{lstlisting}
\caption{
  In this example, the programmer attempts to handle invalid input by
  breaking out of the input loop. When computing the slice, our method
  would record a control dependency between the condition on line 3 and the
  print statement on line 5, placing the bad condition in the program
  slice.
}
\label{early-break}
\end{figure}

@@

>>>>

\subsubsection{Slice (Dynamic)}

\begin{figure}
\begin{lstlisting}
x = input() # Input: 42
if x < 20:
  y = 5
  z = 123
else:
  y = 0
  z = 321
assert(y != 0)
\end{lstlisting}
\begin{lstlisting}
x = 42
if x < 20:

else:
  y = 0

assert(y != 0)
\end{lstlisting}
\caption{
  a) A program being sliced. An assertion gets triggered when line 8 is run. b)
  The lines included in the slice. This is not a valid python
  program, but that does not pose a problem for using membership in the slice as
  a learning feature.
}
\label{fig-slice-example}
\end{figure}

\begin{figure}
\begin{lstlisting}
x = input() # The user inputs 0
if x != 0:
  x = 1
print("One over your number is: %d" % (1 / x))
\end{lstlisting}
\caption{In this program, the programmer may have intended the placeholder
  assignment \lt{x = 1} to be used in case the user input 0, and used the
  incorrect condition \lt{x != 0}, rather than \lt{x == 0}. However, the
  condition expression will not be included in the slice (which would include
  lines 1 and 4), because the presence of
  the conditional does not affect the behavior of the program.
}
\label{fig-slice-downside-example}
\end{figure}

\begin{figure}
\begin{lstlisting}
while true:
  x = input()
  if x != 0:
    break
  print("One over your number is: %d" % (1 / x))
\end{lstlisting}
\caption{
  In this example, the programmer attempts to handle invalid input by
  breaking out of the input loop. When computing the slice, our method
  would record a control dependency between the condition on line 3 and the
  print statement on line 5, placing the bad condition in the program
  slice.
}
\label{fig-early-break}
\end{figure}

@@

\figref{slice-example} is an example of the input and output of the slicing
algorithm. The slicer takes as input a python source string and a list of inputs,
to be fed into the program whenever \lt{input()} is invoked. The slicer outputs
the list of line numbers for the lines in the slice.

The slicer would first create an execution trace of the example program, which
at each execution step includes the line number being run and the state of the
heap and variables. In the example, lines $1, 2, 3, 4, 8$ would have been run.
It then iterates forward through the trace, recording which
execution steps have a define-use relation, and which execution steps
have a test-control relation. The define-use relation is between a step where
the value of a variable is used, and the step where that variable was last
defined. In \figref{slice-example}, a define-use relationship exists
between lines 1 and 2 and between 3 and 8. A test-control relation is between
the execution of a control flow statement and any statements that were run as
a result of it. In \figref{slice-example}, a test-control relationship exists
between the if statement at line 2 and the assignments on lines 3 and 4. After
building up these relations, the slicer starts at the execution step causing an
exception (the execution of line 8), and builds the exception's set of
dependencies by iteratively adding to the set any step that has a define-use
or test-control relationship with a step already in the set. Finally, the
algorithm takes all the execution steps, determines which line number each step
was running, and then outputs this list of line numbers.

Our slicing algorithm handles every syntax node supported by the Python3 standard
library, \lt{ast}, except:
\begin{itemize}
\item Assignments where the left hand side is not a variable or a simple chain
  of attribute indexing or subscripting
\item Assignments where attribute indexing or subscripting on the left hand side
  means something other than the default, which can happen if the operations are
  overridden by a class
\item Lambda expressions
\item Set and dictionary comprehensions
\item Generator expressions
\item Await, yield, and yield from
\item Ellipsis (the token representing the tail of the argument list given to a
  varargs function)
\item Starred expressions
\item Async def functions and Asynchronous for
\item Delete statements
\item With statements
\item Raise statements
\end{itemize}
